---
output:
  bookdown::pdf_document2:
    toc: false
    keep_tex: false
    fig_caption: yes
title: "Title"
author: Author
bibliography: "references.bib"
csl: apa.csl # by default: Chicago style
header-includes:
- \usepackage{float}
- \usepackage{multirow}
- \usepackage{xcolor} 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.lp="fig:")

# install.packages("tinytex", repos = "https://cloud.r-project.org" ) if debug needed?

# library(namelibrary) for the R library

# http://rmarkdown.rstudio.com
# echo = TRUE <- makes all next cmds appear

```
\newcommand{\rocco}[1]{\textcolor{red}{{Rocco:}#1}} <!-- <- this is a command to appear in pdf -->
<!-- command not to appear in pdf -->

# Introduction and Questions

@zhang2004residential cites and refers to @schelling1978micromotives for his prominent influence on threshold and tipping model, still then elaborates on a model based on random utility with single-peaked function and $\beta$. How much can a rational model which uses peak of utility and probability of choice be considered to rely on Schelling's threshold model?
What is the relation between random utility models and threshold models?

```{r chunk1, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}
#echo=TRUE: the code appears in document
#include=TRUE: the output appears in document

```

```{r chunk2, out.widh = '100%', fig.aign='center', fig.cap="CAPTION", include=FALSE}
# knitr::include_graphics("name.pdf")
```



# References
